### 2020-11-25 written/SC_model/SC_model.tex
latest updates

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figs/fig4.pdf}
\end{center}
\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
A. Rapid task switching behavioral paradigm (see text). 
B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
C. The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$. Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.  
D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
\label{fig:SC}
\end{figure}

@@

We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
There is strong pairwise correlation between each parameter of the connectivity matrix.
Shading the distribution by $p_P$, we see that positive shifts in the parameters result in lower accuracy in the Pro task.
Indeed, this difference in accuracy is apparent when we simulate this model for greater parameter values (black star) and lower parameter values (gray star) (Fig. \ref{fig:SC}D).

To make sense of this inferred distributions, we followed the approach of Duan et al. by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig. SX). 
These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair. 
The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$. 
We found that $\lambda_{\text{all}}$ is anti-correlated with $p_P$ and $\lambda_{\text{task}}$ is correlated with $p_P$.
TODO interpret further.  Maybe push further on the optogenetics analysis.

\section{Supplementary}\label{methods_SC}
\begin{equation}
\mathbf{x}_\alpha = f(\mathbf{u}) = \left(\frac{1}{2}\tanh\left(\frac{\mathbf{u}_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
\end{equation}
where $\epsilon = 0.05$ and $\zeta = 0.5$.

The accuracies of $p_P$ and $p_A$ are calculated as
\begin{equation}
p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{x \sim p(x \mid z)}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
\end{equation}
and 
\begin{equation}
p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{x \sim p(x \mid z)}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
\end{equation}
given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function.

@@

>>>>

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figs/fig4.pdf}
\end{center}
\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
A. Rapid task switching behavioral paradigm (see text). 
B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
C. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
D. Simulations of SC networks with parameterizations indicated by the stars in panel C.  Shading  shows standard deviations across trials.
E. The effect of delay period inactivation on accuracy versus Schur eigenvalues.
F. Same as D with delay period inactivation.}
\label{fig:SC}
\end{figure}

@@

We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
To make sense of this inferred distribution, we followed the approach of Duan et al. by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig. \ref{fig:SC_Eigs}A). 
These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair. 
The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$.

In agreement with experimental results form Duan et al., we found the optogenetic silencing during the delay period consistently decreased performance in the Anti task by 15-30\%, but had no consistent effect on the Pro task (Fig. \ref{fig:SC}E). 
For example, the networks indicated by the black star suffers a steep drop in accuracy in the Anti task with delay period inactivation (Fig. \ref{fig:SC}F).  
Networks resilient to delay period inactivation had greater $\lambda_{\text{task}}$ and $\lambda_{\text{task}}$ and lower $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ (Fig. \ref{fig:SC}E).

\section{Supplementary}\label{methods_SC}
\begin{equation}
\mathbf{x}_\alpha = f(\mathbf{u}) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
\end{equation}
where $\epsilon = 0.05$ and $\zeta = 0.5$.

The accuracies of $p_P$ and $p_A$ are calculated as
\begin{equation}
p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
\end{equation}
and 
\begin{equation}
p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
\end{equation}
given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.9]{figs/figSX1.pdf}
\end{center}
\caption{\small Sentence 1 .
A. The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$. Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.  
B. Task accuracy vs. Schur eigenvalue.}
\label{fig:SC_Eigs}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{figs/figSX2.pdf}
\end{center}
\caption{\small
Same as Fig. 1C shaded by Anti task accuracy.}
\label{fig:SC_EPI2}
\end{figure}

### 2020-11-25 written/V1_model/V1_model.tex
latest updates

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{graphicx}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

@@

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Input/output properties of neuron-type population model of V1} \\
Sean Bittner \\
May 12, 2020 \\
\end{center}

\section{Introduction}
In section 3.3 of our bioRxiv preprint \cite{bittner2019interrogating}, we analyze the input/output properties of a 4 neuron-type model of V1.  
We identify degeneracies and sensitivities of neuron-type responses w.r.t specific inputs by visual inspection of the EPI distributions.  
In my opinion, these ``novel insights" are not particularly interesting.
Furthermore, these insights are easily obtained using an existing method -- approximate Bayesian computation (ABC).

In this write-up, I present a new line of inquiry in the same V1 circuit model as the paper.
Rather than conditioning on levels of increase of each neuron-type, we condition on the stabilization of each neuron-type.
EPI then presents us with a distribution of stimulation to this V1 circuit that does not change much a particular neuron-type response.

First, I show EPI converges to a distribution covering the set of samples produced by ABC.
This important sanity check, along with some ABC bias visualization, will make nice appendix figures.
Then, I show how EPI readily identifies the most efficacious patterns of stimulation.
Finally, we confirm two distinct mechanisms of E-population stabilization using EPI on the V1 model with inactivated populations.

\section{V1 model (same as bioRxiv post)}
We consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$ (Fig. 1A). 
Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:

\begin{equation}
\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
\end{equation}

\begin{center}
\textbf{Figure 1: V1 model} \\
\includegraphics[scale=.4]{figs/V1_responses.pdf}
\end{center}

We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic}.
The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.

We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
Maintaining the steady state of a neuron-type population amounts to the emergent property 
\begin{equation}
\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.
\end{equation}
In the following analyses, we chose $\sigma=0.25$.

@@

 \textbf{Figure 2}: EPI (left) vs ABC (right).  Arrows in EPI distribution indicate dimensions of maximal sensitivity at selected parameters $dh$. The importance of $v_S$ and $v_P$ are explained in section 4.2. \\
\begin{center}
$q_\theta(dh \mid \mathcal{B}(E,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(E,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_E.pdf} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_E.png} \\
\clearpage
$q_\theta(dh \mid \mathcal{B}(P,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(P,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_P.png} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_P.png} \\
$q_\theta(dh \mid \mathcal{B}(S,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(S,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_S.png} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_S.png} \\
$q_\theta(dh \mid \mathcal{B}(V,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(V,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_V.png} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_V.png} \\
\end{center}

@@

In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4). 
The regions of the ABC pairplots which have $|dx_\alpha| \approx 0.5$ are far less probable in the EPI distributions (by design), and can account for most mismatch observed.
As a note, there is a discontinuity in Figure 2, $EPI(E, \sigma=0.25)$ in the $dh_E$-$dh_S$ plane.
I'd like to see if I can run the optimization for more random seeds so that we obtain a more connected distribution.

\begin{center}
\textbf{Figure 3: Biased ABC} \\
\includegraphics[scale=0.4]{figs/V1_ABC_Tx.pdf}
\end{center}

\clearpage

\begin{center}
\textbf{Figure 4: Unbiased EPI} \\
\includegraphics[scale=0.4]{figs/V1_EPI_Tx.pdf}
\end{center}

\section{EPI-enabled analyses}
\subsection{Dimensions of sensitivity}
Each arrow in the EPI distributions (Fig. 2, left) points in the direction of maximal sensitivity indicated by the hessian $\partial^2 \log q_\theta(dh) / \partial dh \partial dh^\top$ evaluated at the $dh$ at the base of the arrow.
These patterns of input elicit a greater $dx_\alpha^2$ than (at least) 98.8\% of random stimulation patterns.
Each sensitivity vector is representative of a different stabilization mechanism.

\subsection{Confiriming distinct mechanisms of stabilization}
The sensitivity arrows on the EPI distribution for E-population stability (Fig. 2, top, left) indicate two distinct stability mechanisms for increased $dh_E$.
In one case, $dh_P$ must increase along with increased $dh_E$ (Fig. 2, top, left black arrow $v_P$).
In the other case, $dh_S$ must increase along with increased $dh_E$ (Fig. 2, top, left gray arrow $v_S$).
These inferred distributions and sensitivity vectors $v_P$ and $v_S$ motivate the following hypothesis:

H: If the P- and S-populations are inactivated, the only E-stabilization mechanism with increasing $dh_E$ will be that of the other population (S- or P-population, respectively) .

We can test this hypothesis with EPI, since it is optimized to identify the distribution of all inputs resulting in E-population stability.
We ran EPI with the same emergent property as Eq 2 ($\alpha=E$, $\sigma=0.25$) for two different inactivation scenarios.
In the P-inactivated case, we changed the baseline input to $b =b_{\not P} \triangleq \left[ 1, -5, 1 ,1.25 \right]^\top$, and to $b =b_{\not S} \triangleq \left[ 1, 1, -5 ,1.25 \right]^\top$ in the S-inactivated case.
To test for the existence of stabilization mechanisms corresponding to $v_P$ and $v_S$, we measured the most-sensitive dimension of the Hessian at all points in the inferred distribution.  
We declare that the P- or S-inactivated circuit lacks a stability mechanism, if no sensitivity vectors throughout the distribution are aligned with the corresponding sensitivity vector.

\begin{center}
\textbf{Figure 5}: Stability mechanisms of inactivated circuits.
\includegraphics[scale=.5]{figs/cos_dists.pdf}
\end{center}

Confirming our hypothesis, we found that only the $v_S$ stability mechanism is present in the P-inactivated circuit (Figure 5A), and that only the $v_P$ stability mechanism is present in the S-inactivated circuit (Figure 5B).
As expected, the samples of the inferred distributions aligned with the stability mechanism correspond to the sloped regions orthogonal to $v_S$ and $v_P$ (Figure 6).

\clearpage

\begin{center}
\textbf{Figure 6}: Inactivated circuit stability distributions.
\includegraphics[scale=.4]{figs/EPI_inactivation.pdf}
\end{center}

\section{Summary}
We have improved on our previous EPI analysis of V1 in the following ways:
\begin{enumerate}
\item We utilize a novel ability of EPI (Hessian evaluation).
\item Our scientific conclusion is about mechanisms of stability rather than high-level characteristics of responses.
\item We have validation of structure via ABC.
\item We characterize the issue of bias in ABC (and lack thereof in EPI).
\item The emergent property of ``neuron-type stabilization" follows the ethos of the definition of emergent property better than ``neuron-type population rate increase of $y$".
\end{enumerate}

I recommend substituting this work for section 3.3 in the bioRxiv manuscript. I'd be happy to write up a draft of such a section.

%\section{Isolated modulation paths using EPI}
%Let's say we want to find a stimulation pattern that changes the $\alpha$-population steady-state response $dx_\alpha$ as much as possible while keeping the rest the same.
%One can not simply stimulate the $\alpha$-population, since one must take into account the effects of the recurrent connectivity.
%Furthermore, such an input pattern producing isolated modulation may change nonlinearly.
%
%Beginning at $dh = \textbf{0}$, we seek a sequence of differential inputs that will increasingly change $dx_\alpha$, while maintaining $dx_\beta$, $\beta \in \{E, P, S, V\} \setminus	\{\alpha\}$.
%At each step in the path of inputs, the direction is selected to maximize change in $dx_\alpha$ and minimize change in $dx_{\beta \neq \alpha}$.
%These input selections can be guided by the EPI distributions $q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ that result in the modulation of the $\alpha$ population within a given range.
%
%We leverage the availability of a functional representation of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ to calculate the the hessians of the EPI distributions at each step of $dh$ in the path
%\begin{equation}
%\text{Hess}_\alpha(dh) = \frac{\partial^2 \log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))}{\partial dh \partial dh^\top}.
%\end{equation}
%
% We select the optimal direction of isolated modulation as vector v, which points in the maximally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$, and minimally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\beta, 0.2))$, $\beta \neq \alpha$:
% \begin{equation}
% v_\alpha^*(dh) = \argmin_v v^\top \text{Hess}_\alpha(dh) v - \sum_{\beta \neq \alpha} v^\top \text{Hess}_\beta(dh) v,
% \end{equation}
%  where $v\top v$ = 1.  $v_\alpha(dh)$ is obtained via an eigendecomposition:
%  \begin{equation}
%  v_\alpha^*(dh) = \argmin_v v^\top M_\alpha(dh) v
% \end{equation}
% \begin{equation}
%M_\alpha(dh) = \text{Hess}_\alpha(dh) - \sum_{\beta \neq \alpha} \text{Hess}_\beta(dh).
% \end{equation}
% The solution is the eigenvector with smallest eigenvalue of $M_\alpha(dh)$.
% 
% \includegraphics[scale=0.4]{figs/hess_paths.pdf}
%

>>>>

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{systeme}
\usepackage{graphicx}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\usepackage{bm}

@@

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Draft of new V1 section} \\
Sean Bittner, Agostina Palmigiano \\
November 24, 2020 \\
\end{center}

\section{Exploratory analysis of V1 with EPI produces a novel theory}\label{results_V1}
Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}. 
Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig. \ref{fig:SV_flip}A) \cite{pfeffer2013inhibition}.
Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools. 
Here, we use EPI to elucidate the mechanisms of neuron-type stability in V1 at different levels of contrast.

We consider contrast responses of a nonlinear dynamical V1 circuit model (Fig. \ref{fig:Fig3}A) with a state comprised of each neuron-type population's rate $x = \left[x_E, x_P , x_S, x_V \right]^\top$.
Each population of neuron-type $\alpha$ receives recurrent input $\left(W f_r(x)\right)_\alpha$ from synaptic projections, where $W$ is the connectivity matrix and $f_r = \left[\cdot\right]_+^2$ is the rate nonlinearity.
Driven by a baseline input $\mathbf{b}$, the circuit evolves form an initial condition $\mathbf{x}(0)$ to a steady-state solution (Fig. \ref{fig:Fig3}B, solid).
When slow noise $\mathbf{\epsilon}$ is introduced, circuit activity fluctuates around this solution (Fig. \ref{fig:Fig3}B dashed), and the model is then a stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} (see Section \ref{methods_V1}).
As contrast is enhanced, input to the E- and P-populations via $h^{(c)}$ increases causing the steady state solution (and fluctuations thereabout) to change  (Fig. \ref{fig:Fig3}C).
In this analysis, we consider $W$,$b$, and $h^{(c)}$ that have been fit to contrast responses in mouse V1 using the deterministic model (TODO cite, see Section \ref{methods_V1}).
As contrast changes, so does the degree of stochastic variability of each neuron-type (Fig. \ref{fig:Fig3}D).
Moreover, the circuit transitions into an inhibition-stabilized network (ISN) with increasing contrast (Fig. \ref{fig:Fig3}D inset).
While the ISN-regime and it's effects are well understood in E-I networks, it's role in the stability of this SSSN with inhibitory multiplicity has not been explored.

To study the role of inhibition stabilization in this network, we explore the changes in input 
\begin{equation}
\mathbf{z} = \begin{bmatrix} dh_E & dh_P & dh_S & dh_V \end{bmatrix}^{\top}
\end{equation}
resulting in each neuron-type's stability around it's mean rate.
The emergent property of ``neuron-type stability" for population $\alpha$  at contrast level $c$ is defined as 
\begin{equation}\label{eq:EP}
\begin{split}
\mathcal{X}(\alpha, c) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  0  ~~\triangleq~~ \bm{\mu}  \\ 
\text{Var}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0.05^2 \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
\end{split}
\end{equation}

\clearpage
\begin{figure}[h]
\caption{\small \textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.   Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).
\textbf{B}. EPI posterior $q_\theta(z \mid \mathcal{B}_{S-V})$ for S-V flipping. The obtained posterior is visualized as 500 samples from the inferred distribution colored by $\log(q_\theta(z))$. This posterior is bimodal and concentrated in planes $h^{(c)} > 0$ and $h^{(c)} < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.  Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two standard deviations). 
\textbf{C}. Posterior predictive distribution of $d_{S,V}$ for each mode.  The $z_1$-mode produces V-to-S flipping with increasing contrast.
\textbf{D}. Model simulations at the mode $z_1$ at $c=0$ (solid) and $c=1$ (dashed). \textbf{E}. Mean system response at varying contrasts for $z_1$. Error bars show noise in rate. \textbf{F}. Plot of noise with contrast from E. Error bars are standard error measured across simulations.
 }\label{fig:Fig3}
\begin{center}
\includegraphics[scale=.18]{figs/Fig3/Fig3.pdf}
\end{center}
\end{figure}
\clearpage

\section{V1 Supplemental section}\label{methods_V1}

The dynamics are driven by the rectified and exponentiated sum of recurrent inputs $Wx$, external inputs $h$, and external noise:

\begin{equation}
\tau \frac{dx}{dt} = -x +Wf_r(x) + b + h + \epsilon
\end{equation}

where $\tau = 1 \text{ms}$, and $f_r(x) = \left[x \right]_+^2$.

The noise is modeled as an Ornstein-Uhlenbeck process:
\begin{equation}
\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB
\end{equation}

where $\tau_{\text{noise}}$ is slow at 5ms, $\sigma_\alpha =  0.01 $, and $dB$ is a standard Wiener process.

We considered the stability of this stochastic stabilized supralinear network using connectivity ($W$) and input ($b$ and $h^{(c)}$) parameters fit to the deterministic model (TODO cite)

\begin{equation}
W =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
 \begin{bmatrix} .785 & -.102 & -1.24 & -.306 \\
.809 & -.100 & -.620 & -.264 \\
.833 & -.000128 & -.0000483 & -.742 \\
.708 & -.306 & -.452 & -.0000605 \\
 \end{bmatrix},
\end{equation}

\begin{equation}
b =  \begin{bmatrix} b_E \\ b_P \\ b_S \\ b_V \end{bmatrix} =
 \begin{bmatrix} .590 \\ .504 \\ .515 \\ .670 \end{bmatrix} ,
\end{equation} 
and
\begin{equation} 
h^{(c)} = \begin{bmatrix} h^{(c)}_E \\ h^{(c)}_P \\ 0 \\ 0 \end{bmatrix} = 
\begin{bmatrix} .595 \\ .396 \\ 0 \\ 0 \end{bmatrix}.
\end{equation}

This ISN coefficient $\gamma(x; z)$ is calculated by simulating

The expectation and variance in the emergent property definition are taken over both the posterior distribution and the stochasticity of the SSSN:
\begin{equation}
= \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  = \mathbb{E}_{\mathbf{z}\sim q_\theta} \left[ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \begin{bmatrix} x_\alpha(\mathbf{x}; \mathbf{z},c) - x_\alpha(\mathbf{x}; \mathbf{0},c) \end{bmatrix}\right].
\end{equation}

\begin{figure}[h]
\caption{\small SX 1
 }\label{fig:SX1}
\begin{center}
\includegraphics[scale=.5]{figs/FigSX/Fig_SX1.pdf}
\end{center}
\end{figure}

\begin{figure}[h]
\caption{\small SX 2
 }\label{fig:SX2}
\begin{center}
\includegraphics[scale=.5]{figs/FigSX/Fig_SX2.pdf}
\end{center}
\end{figure}

 With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
 Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
 All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
 While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig. \ref{fig:V1_EPI}B right).  
 Currently available approaches to deriving the steady state response of the system are limited.

%\section{Stability analysis}
The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.

We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
Maintaining the steady state of a neuron-type population amounts to the emergent property 
\begin{equation}
\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.\\
\end{equation}
In the following analyses, we chose $\sigma=0.25$.

@@

 \textbf{Figure 2}: EPI (left) vs ABC (right).  Arrows in EPI distribution indicate dimensions of maximal sensitivity at selected parameters $dh$. The importance of $v_S$ and $v_P$ are explained in section 4.2. \\

@@

In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4).
